\documentclass[12pt, a4paper]{report}

\usepackage[refpage]{nomencl}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{setspace}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{rotating}
\usepackage{gensymb} 
\usepackage{tabularx}
\usepackage{subcaption}
\usepackage[hidelinks]{hyperref}
\usepackage{color}
\usepackage[toc,page]{appendix}
\usepackage{tikz}

\definecolor{sh_comment}{rgb}{0.12, 0.38, 0.18 } %adjusted, in Eclipse: {0.25, 0.42, 0.30 } = #3F6A4D
\definecolor{sh_keyword}{rgb}{0.37, 0.08, 0.25}  % #5F1441
\definecolor{sh_string}{rgb}{0.06, 0.10, 0.98} % #101AF9

\DeclareCaptionFont{white}{\color{white}}
\DeclareCaptionFormat{listing}{%
  \parbox{\textwidth}{\colorbox{gray}{\parbox{\textwidth}{#1#2#3}}\vskip-4pt}}
\captionsetup[lstlisting]{format=listing,labelfont=white,textfont=white}
% \lstset{frame=lrb, xleftmargin=\fboxsep, xrightmargin=-\fboxsep, breaklines=true, basicstyle=\ttfamily\scriptsize}
\lstset {
 frame=ltrb,
 rulesepcolor=\color{black},
 showspaces=false,showtabs=false,tabsize=2,
 numberstyle=\tiny,numbers=left,
 basicstyle= \ttfamily\scriptsize,
 stringstyle=\color{sh_string},
 keywordstyle = \color{sh_keyword}\bfseries,
 commentstyle=\color{sh_comment}\itshape,
 captionpos=b,
 xleftmargin=0.7cm, xrightmargin=0.5cm,
 lineskip=-0.3em,
 float,
 floatplacement=H,
 escapebegin={\lstsmallmath}, escapeend={\lstsmallmathend},
 breaklines=true
}

\onehalfspacing

\makenomenclature
 
\begin{document}

\title{Using Local and Global Knowledge in Wireless Sensor Networks}
\author{Christopher Gwilliams\\Cardiff University}
\date{\today}
\maketitle 
\pagenumbering{roman}
\tableofcontents
\listoffigures
\listoftables

\begin{abstract}
Wireless sensor networks (WSNs) have advanced rapidly in recent years and the volume of raw data received at an endpoint can be huge. We believe that the use of local knowledge, acquired from sources such as the surrounding environment, users and previously sensed data, can improve the efficiency of a WSN and automate the classification of sensed data. We define local knowledge as knowledge about an area that has been gained through experience or experimentation. With this in mind, we have developed a three-tiered architecture for WSNs that uses differing knowledge-processing capabilities at each tier, called the Knowledge-based Hierarchical Architecture for Sensing (K-HAS). A novel aligning ontology has been created to support K-HAS, joining widely used, domain-specific ontologies from the sensing and observation domains. We have shown that, as knowledge-processing capabilities are pushed further out into the network, the profit - defined as the value of sensed data - is increased; where the profit is defined as the value of the sensed data received by the end user.

Collaborating with Cardiff University School of Biosciences, we have deployed a variation of K-HAS in the Malaysian rainforest to capture images of endangered wildlife, as well as to automate the collection and classification of these images. Technological limitations prevented a complete implementation of K-HAS and an amalgamation of tiers was made to create the Local knowledge Ontology-based Remote-sensing Informatics System (LORIS). A two week deployment in Malaysia suggested that the architecture was viable and that, even using local knowledge at the endpoint of a WSN, improved the efficiency of the network. A simulation was implemented to model K-HAS and this indicated that the network became more efficient as knowledge was pushed further out towards the edge, by allowing nodes to prioritise sensed data based on inferences about its content.
\end{abstract}

\pagenumbering{arabic}

\include{Chap1/introduction}
\include{Chap2/background}
\include{Chap3/technical}
\include{Chap4/architecture}
\include{Chap5/ont}
\include{Chap6/implementation}
\include{Chap7/simulations}
\include{Chap8/conclusion}


\bibliographystyle{plain}
\bibliography{refs/library}
\include{App/appendix}
\end{document}
