% This thesis template has been put together and edited by Diego Pizzocaro
% It's a collection/edit of files produced by Frank C Langbein for his PhD thesis
% and then edited by Ahmed Alazzawi, Hmood Al Dossari, Manar I. Hosny, H.Z. Al-Dossari 
% and kindly shared by them.
%
% Following, the original GNU license chosen by Frank C Langbein, 
% which is the same license for distributing this package.
%
% beautification.tex -- Beautification of Reverse Engineered Geometric Models
%% PhD Thesis by Frank C. Langbein%
% Maintainer: Frank C. Langbein <frank@langbein.org>
% Version: 1.0
% Copyright (c) 2003 Frank C Langbein.
% Permission is granted to copy, distribute and/or modify this document
% under the terms of the GNU Free Documentation License, Version 1.2 or
% any later version published by the Free Software Foundation; with no
% Invariant Sections, no Front-Cover Texts, and no Back-Cover Texts. A
% copy of the license is included in the section entitled "GNU Free
% Documentation License".
%


\documentclass[a4paper,oneside,onecolumn,openright,12pt]{book}


\makeatletter

%
\usepackage{caption}
\usepackage{footmisc} 
\usepackage{paralist}
\usepackage{amsthm}
% \usepackage[utf8]{inputenc}
% \usepackage{subfig}
%\usepackage{algorithm}
%\usepackage{graphicx}
%\usepackage{rotating}
%\usepackage{amssymb} 


% Fonts, encoding, etc.
\usepackage{type1cm}
% \usepackage[latin1]{inputenc}
\usepackage[british]{babel}
\usepackage[T1]{fontenc}
\usepackage{times}
% \usepackage{hyperref}
\usepackage{longtable}
%\usepackage{algorithm} 
\usepackage[noend]{algorithmic}
%\usepackage{graphicx}
%\usepackage{rotating}
\usepackage{multirow}
\usepackage{float}


\usepackage[numbers]{natbib}

\usepackage[refpage]{nomencl} 
\usepackage{graphicx}
\usepackage{listings}
\usepackage{setspace}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{rotating}
\usepackage{gensymb} 
\usepackage{tabularx}
\usepackage{subcaption}
\usepackage[hidelinks]{hyperref}
\usepackage[all]{hypcap}
\usepackage{color}
\usepackage[toc,page]{appendix}
\usepackage{tikz}

\definecolor{sh_comment}{rgb}{0.12, 0.38, 0.18 } %adjusted, in Eclipse: {0.25, 0.42, 0.30 } = #3F6A4D
\definecolor{sh_keyword}{rgb}{0.37, 0.08, 0.25}  % #5F1441
\definecolor{sh_string}{rgb}{0.06, 0.10, 0.98} % #101AF9

\DeclareCaptionFont{white}{\color{white}}
\DeclareCaptionFormat{listing}{%
  \parbox{\textwidth}{\colorbox{gray}{\parbox{\textwidth}{#1#2#3}}\vskip-4pt}}
\captionsetup[lstlisting]{format=listing,labelfont=white,textfont=white}
% \lstset{frame=lrb, xleftmargin=\fboxsep, xrightmargin=-\fboxsep, breaklines=true, basicstyle=\ttfamily\scriptsize}
\lstset {
 frame=ltrb,
 rulesepcolor=\color{black},
 showspaces=false,showtabs=false,tabsize=2,
 numberstyle=\tiny,numbers=left,
 basicstyle= \ttfamily\scriptsize,
 stringstyle=\color{sh_string},
 keywordstyle = \color{sh_keyword}\bfseries,
 commentstyle=\color{sh_comment}\itshape,
 captionpos=b,
 xleftmargin=0.7cm, xrightmargin=0.5cm,
 lineskip=-0.3em,
 float,
 floatplacement=H,
 escapebegin={\lstsmallmath}, escapeend={\lstsmallmathend},
 breaklines=true
}

%\usepackage{subfigure}
% Line spacing
\def\baselinestretch{1.5}
\parindent0cm
\parskip1.5ex\@plus.7ex\@minus.1ex\relax

%Acronyms
 \usepackage{acronym}
 %Inline lists
 \usepackage{paralist}

% Page dimensions
\usepackage{vmargin}
\setpapersize{A4}
\setmarginsrb{40mm}{20mm}{25mm}{30mm}{14.5pt}{8mm}{0pt}{11mm}

% Footer and header
\usepackage{afterpage}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhead{}
\fancyhead[LE,RO]{\thepage}
\fancyhead[LO,RE]{\slshape \leftmark}
\fancyfoot{}
\renewcommand{\chaptermark}[1]{}
\renewcommand{\sectionmark}[1]%
             {\markboth{\thesection\ #1}{\thesection\ #1}}
\renewcommand{\subsectionmark}[1]{}
\fancypagestyle{plain}{%
  \fancyhead{}
  \fancyhead[LE,RO]{\thepage}
  \fancyfoot{}
  \renewcommand{\headrulewidth}{.6pt}
}

% Chapter
\def\@makechapterhead#1{%
  \ \\[-35.5pt]\hbox to \textwidth {%
    \hfill {\vbox{\hbox{\rule[5pt]{140pt}{4pt}}%
        \hbox to 140pt {\hfill\huge\bfseries\slshape \@chapapp\space\thechapter\/}}}}%
  \vskip55\p@%
  {\parindent \z@ \raggedright \normalfont%
    \interlinepenalty\@M%
    \Huge \bfseries #1\par\nobreak%
    \vskip 45\p@%
  }}
\def\@chapter[#1]#2{\ifnum \c@secnumdepth >\m@ne
                       \if@mainmatter
                         \refstepcounter{chapter}%
                         \typeout{\@chapapp\space\thechapter.}%
                         \addcontentsline{toc}{chapter}%
                                   {\protect\numberline{\thechapter}#1}%
                       \else
                         \addcontentsline{toc}{chapter}{#1}%
                       \fi
                    \else
                      \addcontentsline{toc}{chapter}{#1}%
                    \fi
                    \chaptermark{#1}%
                    \addtocontents{lof}{\protect\addvspace{10\p@}}%
                    \addtocontents{lot}{\protect\addvspace{10\p@}}%
                    \addtocontents{loa}{\protect\addvspace{10\p@}}%
                    \if@twocolumn
                      \@topnewpage[\@makechapterhead{#2}]%
                    \else
                      \@makechapterhead{#2}%
                      \@afterheading
                    \fi}
\def\@schapter#1{\addcontentsline{toc}{chapter}{#1}%
                 \markboth{#1}{#1}%
                 \addtocontents{lof}{\protect\addvspace{10\p@}}%
                 \addtocontents{lot}{\protect\addvspace{10\p@}}%
                 \addtocontents{loa}{\protect\addvspace{10\p@}}%
                 \if@twocolumn%
                    \@topnewpage[\@makeschapterhead{#1}]%
                 \else%
                    \@makeschapterhead{#1}%
                    \@afterheading%
                 \fi}
\def\@makeschapterhead#1{%
  \ \\[-35.5pt]\hbox to \textwidth {%
    \hfill {\vbox{\hbox{\rule[5pt]{140pt}{0pt}\rule[5pt]{0pt}{4pt}}%
        \hbox to 140pt {\hfill\huge\bfseries\slshape \ \/}}}}%
  \vskip55\p@%
  {\parindent \z@ \raggedright \normalfont%
    \interlinepenalty\@M%
    \Huge \bfseries  #1\par\nobreak%
    \vskip 45\p@%
  }}

% Table of contents
\def\contentsname{Contents}
\renewcommand\tableofcontents{%
    \if@twocolumn%
      \@restonecoltrue\onecolumn%
    \else%
      \@restonecolfalse%
    \fi%
    \chapter*{\contentsname}%
    \@starttoc{toc}%
    \if@restonecol\twocolumn\fi%
    }

% Bibliography
\renewenvironment{thebibliography}[1]
     {\chapter*{\bibname}%
      \list{\@biblabel{\@arabic\c@enumiv}}%
           {\settowidth\labelwidth{\@biblabel{#1}}%
            \leftmargin\labelwidth
            \advance\leftmargin\labelsep
            \@openbib@code
            \usecounter{enumiv}%
            \let\p@enumiv\@empty
            \renewcommand\theenumiv{\@arabic\c@enumiv}}%
      \sloppy
      \clubpenalty4000
      \@clubpenalty \clubpenalty
      \widowpenalty4000%
      \sfcode`\.\@m}
     {\def\@noitemerr
       {\@latex@warning{Empty `thebibliography' environment}}%
      \endlist}

% Floats
\long\def\@makecaption#1#2{%
  \vskip\abovecaptionskip
  \sbox\@tempboxa{\textbf{#1: #2}}%
  \ifdim \wd\@tempboxa >\hsize
    \textbf{#1: #2.}\par
  \else
    \global \@minipagefalse
    \hb@xt@\hsize{\hfil\box\@tempboxa\hfil}%
  \fi
  \vskip\belowcaptionskip}
\renewcommand{\topfraction}{0.9}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.9}

% Tables
\usepackage{dcolumn}
\usepackage{hhline}

% Graphics
% \usepackage[dvips]{graphicx}
% \usepackage[usenames,dvipsnames]{color}
\usepackage{rotating}
\usepackage{psfrag}
\usepackage{epic}
\usepackage{eepic}

% Algorithm environment
\newcounter{algorithm}[chapter]
\renewcommand{\thealgorithm}{\thechapter.\@arabic\c@algorithm}
\def\fps@algorithm{t}
\def\ftype@algorithm{1}
\def\ext@algorithm{loa}
\def\fnum@algorithm{Algorithm~\thealgorithm}
\newenvironment{algorithm}{\@float{algorithm}}{\end@float}
\newenvironment{algorithm*}{\@dblfloat{algorithm}}{\end@dblfloat}
\newenvironment{alevel}%
   {\begin{list}{}{%
      \setlength{\topsep}{0pt}%
      \setlength{\parskip}{0pt}%
      \setlength{\partopsep}{0pt}%
      \setlength{\parsep}{0pt}%
      \setlength{\itemsep}{0pt}}}%
   {\end{list}}
\newcommand\listalgorithmsname{List of Algorithms}
\newcommand\listofalgorithms{%
    \if@twocolumn
      \@restonecoltrue\onecolumn
    \else
      \@restonecolfalse
    \fi
    \chapter*{\listalgorithmsname}%
      \@mkboth{\listalgorithmsname}{\listalgorithmsname}%
    \@starttoc{loa}%
    \if@restonecol\twocolumn\fi
    }
\newcommand*\l@algorithm{\@dottedtocline{1}{1.5em}{2.3em}}

% List of figures and tables
\renewcommand\listoffigures{%
    \if@twocolumn
      \@restonecoltrue\onecolumn
    \else
      \@restonecolfalse
    \fi
    \chapter*{\listfigurename}%
      \@mkboth{\listfigurename}{\listfigurename}%
    \@starttoc{lof}%
    \if@restonecol\twocolumn\fi
    }
\renewcommand\listoftables{%
    \if@twocolumn
      \@restonecoltrue\onecolumn
    \else
      \@restonecolfalse
    \fi
    \chapter*{\listtablename}%
      \@mkboth{\listtablename}{\listtablename}%
    \@starttoc{lot}%
    \if@restonecol\twocolumn\fi
    }

% Math symbols, fonts, etc.
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}

\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\renewcommand{\S}{\mathbb{S}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}

\newcommand{\Cf}{\mathfrak{C}}
\newcommand{\Pf}{\mathfrak{P}}

\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\avg}{avg}
\DeclareMathOperator{\floor}{floor}
\DeclareMathOperator{\ceil}{ceil}
\DeclareMathOperator{\round}{round}

\providecommand{\abs}[1]{\lvert#1\rvert}
\providecommand{\absd}[1]{\left\lvert#1\right\rvert}
\providecommand{\card}[1]{\lvert#1\rvert}
\providecommand{\norm}[1]{\lVert#1\rVert}

% URLs
\usepackage{url}
%% Define a new 'leo' style for the package that will use a smaller font.
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\footnotesize\ttfamily}}}
\makeatother
%% Now actually use the newly defined style.
\urlstyle{leo}



\makeatother


\begin{document}


\frontmatter


\begin{titlepage}

\begin{center}
\vspace*{3ex}
\textbf{\Huge Using Local and Global Knowledge}\\[2ex]
\textbf{\Huge in Wireless Sensor Networks}\\[12ex]
\textbf{\large A thesis submitted in partial fulfilment}\\[1ex]
\textbf{\large of the requirement for the degree of Doctor of
  Philosophy}\\[16ex]
\textbf{\LARGE Christopher Gwilliams}\\
\vfill
\textbf{\LARGE \today}\\
\vfill
\textbf{\LARGE Cardiff University}\\[1ex]
\textbf{\LARGE School of Computer Science \& Informatics}\\[4ex]
\end{center}

\end{titlepage}
\newpage\thispagestyle{empty}\cleardoublepage


\thispagestyle{plain}

\vspace*{6ex}

\textbf{\large Declaration}

This work has not previously been accepted in substance for any degree and is not concurrently submitted in candidature for any degree.\\[2ex]
Signed \dotfill \ (candidate) \hspace*{10em}\\[1ex]
Date\ \ \ \ \ \dotfill \hspace*{18em}

\vfill

\textbf{\large Statement 1}

This thesis is being submitted in partial fulfilment of the requirements for the degree of PhD.\\[2ex]
Signed \dotfill \ (candidate) \hspace*{10em}\\[1ex]
Date\ \ \ \ \ \dotfill \hspace*{18em}

\textbf{\large Statement 2}

This thesis is the result of my own independent work/investigation,
except where otherwise stated. Other sources are acknowledged by
explicit references.\\[2ex]
Signed \dotfill \ (candidate) \hspace*{10em}\\[1ex]
Date\ \ \ \ \ \dotfill \hspace*{18em}

\vfill

\textbf{\large Statement 3}

I hereby give consent for my thesis, if accepted, to be available for photocopying and for inter-library loan,
 and for the title and summary to be made available to outside organisations.\\[2ex]
Signed \dotfill \ (candidate) \hspace*{10em}\\[1ex]
Date\ \ \ \ \ \dotfill \hspace*{18em}

\vfill

\cleardoublepage

\thispagestyle{plain}
\ \vfill{\small
Copyright \copyright\ 2014 Christopher Gwilliams.\\
Permission is granted to copy, distribute and/or modify this document
under the terms of the GNU Free Documentation License, Version 1.2 or
any later version published by the Free Software Foundation; with no
Invariant Sections, no Front-Cover Texts, and no Back-Cover Texts. A
copy of the license is included in the section entitled ``GNU Free
Documentation License''.}\\[3.5ex]
A copy of this document in various transparent and opaque
machine-readable formats and related software is available at
\url{http://yourwebsite}.
\cleardoublepage


%\ \vspace*{1.11cm}
%\markboth{Dedication}{}
%\begin{flushright}
%\textbf{\large To People you care}\\
%\textbf{\large for their patience and support.}
%\end{flushright}
%\newpage
%\markboth{}{}
%\cleardoublepage


\chapter*{Abstract}

%\def\baselinestretch{1}\normalfont

\def\baselinestretch{1.5}\normalfont

Wireless sensor networks (WSNs) have advanced rapidly in recent years and the volume of raw data received at an endpoint can be huge. We believe that the use of local knowledge, acquired from sources such as the surrounding environment, users and previously sensed data, can improve the efficiency of a WSN and automate the classification of sensed data. We define local knowledge as knowledge about an area that has been gained through experience or experimentation. With this in mind, we have developed a three-tiered architecture for WSNs that uses differing knowledge-processing capabilities at each tier, called the Knowledge-based Hierarchical Architecture for Sensing (K-HAS). A novel aligning ontology has been created to support K-HAS, joining widely used, domain-specific ontologies from the sensing and observation domains. We have shown that, as knowledge-processing capabilities are pushed further out into the network, the profit - defined as the value of sensed data - is increased; where the profit is defined as the value of the sensed data received by the end user.

Collaborating with Cardiff University School of Biosciences, we have deployed a variation of K-HAS in the Malaysian rainforest to capture images of endangered wildlife, as well as to automate the collection and classification of these images. Technological limitations prevented a complete implementation of K-HAS and an amalgamation of tiers was made to create the Local knowledge Ontology-based Remote-sensing Informatics System (LORIS). A two week deployment in Malaysia suggested that the architecture was viable and that, even using local knowledge at the endpoint of a WSN, improved the efficiency of the network. A simulation was implemented to model K-HAS and this indicated that the network became more efficient as knowledge was pushed further out towards the edge, by allowing nodes to prioritise sensed data based on inferences about its content.

\chapter*{Acknowledgements}

My eternal gratitude goes to Alun Preece and Alex Hardisty for their guidance, assistance, support, enthusiasm and faith in me throughout the research and writing of this thesis. Their encouragement of testing new ideas, questioning everything and ensuring I step out of the box that it seems all too easy to stay inside. I am also grateful for the skills they have pushed me to gain, from teaching to presenting to professional writing. None of this work would have been possible without you both.

Benoit Goosens, Mike Bruford, Danica Stark and all of those involved with Danau Girang Field Centre have been fountains of knowledge, support and inspiration. At the very least, I thank you for your hospitality each year and to the staff members that put up with a computer scientist that is clearly not well adapted to life in the rainforest with limited connectivity to the outside world. 

I am grateful to the School of Computer Science \& Informatics, staff and students have provided methods of growth, new skills and support throughout my time at Cardiff University. My thanks go to Matt Williams, for challenging my work from the early stages right through to completion and providing relief through films, and Will Webberley, a friend gained through the PhD that provided support, both professionally and personally, and I do not believe I could have completed this work without him. Special mentions must go to Martin Chorley, Jonathan Quinn, Gualtiero Colombo, Konrad Borowiecki, Diego Pizzocaro, Przemyslaw Woznowski, Rich Coombs, Liam Turner, Mark Greenwood and Matthew John for being there in the pub, as well as in academia.

My deepest thanks also go to family and close friends, with special mentions to Natasha Kinsley-York for supporting me throughout the start of the PhD and helping me deal with the stress incurred and Chris Fellows, for his assistance and also being a sounding board for my many rants, ideas and everything in between. Finally, to Katie Lewis, for her ideas, support, patience and all-round awesomeness that has not only helped my work, but ensured I enjoy every second of my down time.

I would like to dedicate this work to my father, who passed away in 2009, there is no way I would be anywhere close to this point without his input, teaching and encouragement pushing me to go further than I ever imagined I could go. Words cannot express the love and gratitude I have for Robert and Gina Gwilliams, two role models that truly set the bar high. 



\tableofcontents

\input{listofpublications.tex}

\listoffigures

\listoftables

\lstlistoflistings

% \listofalgorithms

\input{acronyms.tex}

\mainmatter
%begin each chapter chapter1.tex by \Chapter{chapter1}, for example, then the sections and so on...
\include{Chap1/introduction}
\include{Chap2/background}
\include{Chap3/technical}
\include{Chap4/architecture}
\include{Chap5/ont}
\include{Chap6/implementation}
\include{Chap7/simulations}
\include{Chap8/conclusion}

\bibliographystyle{plainnat}
\bibliography{refs/library}
\include{App/appendix}
\backmatter


\end{document}
